Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-08-18 writeup/RochatEtal_ResponseToReviewers.tex

      {\it ``Any horizontal velocity field, uh, can be expressed in terms of vorticity, divergence,
      strain rate and mean flow. This can be seen by expressing uh as a Taylor series
      expansion and then decomposing the velocity gradient into symmetric and anti-
      symmetric parts [Landau and Lifshitz, 1987]. Because vortices, vertical motion and
      fronts are ubiquitous at the submesoscale, this is a natural decomposition and
      variances of these quantities are apt descriptors of submesoscale turbulence.''}\\

      \bdp
        Adding the elementary motivation above is yet another controversy: some
        readers may appreciate it but others may find it condescending.
        We argue that most physical oceanographers interested in our paper will
        likely know this material. Even readers from other areas, who did not take
        basic fluid mechanics, will have the intuition that if the flow peaks
        at mesoscales (energy-containing), then second-order statistics of the
        velocity gradient will better highlight the submesoscales than the
        KE (second-order statistic of the velocity).
      \edp

\item {\it (Figure 3) Would be nice to include a third row that contains the collapsed (i.e., one-
      dimensional) PDFs of normalized vorticity to illustrate the skewness of hourly and
      daily-averaged flow. I believe you already have this information in the supplementary
      material but would be nice to illustrate to the reader. Also, in your normalization of
      vorticity, are you accounting for the fact that the Coriolis parameter changes with
      latitude? Your domain spans considerable changes in f and the relevant dynamical
      parameter is the gradient Rossby number.}

      \bdp
        We have experimented with a new figure 3, including a third row to show the
        vorticity PDFs. But the extra plots add little extra, if any, information
        and take a lot of space. For what it is worth, we explicitly give the relevant statistic
        (skewness) for each month, etc.  As the reviewer mentions, the collapsed
        PDFs for vorticity and divergence are shown in Figure S2, which is a plot
        that compares April/October and hourly/daily-averaged in a single plot.\\

        Yes, as we mention in the text, we have normalized by those
        quantities by the \textit{local} inertial frequency, not the inertial
        frequency at mid-latitude.
      \edp

@@

\item {\it (Figure 4b) Strictly speaking, the label on the y-axis should not be SSH variance. SSH
      variance would be the integral of this curve. The y-axis should be labelled as the
      power spectral density of sea surface height (SSH), or SSH Spectra.}\\

      \bdp
      We labelled the y-axis of the SSH variance spectrum as \textit{SSH variance density}
      --- short for \textit{SSH variance spectral density} --- not \textit{SSH variance} as the reviewer
      suggests.  We chose to use the short term to avoid cluttering the label
      since there is no ambiguity because we include the
      units of SSH variance per unit wavenumber.
      We further disagree that the y-axis should be labelled \textit{power spectral density of SSH}
      or \textit{SSH spectra} [{\it sic}]. On physical grounds, \textit{power} adds
      unnecessary ambiguity,
      because SSH variance density has nothing to do with \textit{power}.\footnote{The
      term \textit{power spectral density}  appears to have been introduced
      for frequency spectrum of various quantities in signal processing (Rob Pinkel, personal
      communication).}
      And ``SSH [wavenumber] spectrum'' is the plot of SSH variance spectral density
      as a function of wavenumber, not the quantity in the y-axis.\footnote{This confusion is similar
      to the pervasive axis mislabelling in plots of probability distribution function (PDF)
      --- the best label is \textit{probability density} not PDF.}
      \edp

\item {\it The authors may find the work of Brüggeman and Eden (2015) helpful in addressing
      some of the aforementioned questions.}\\

      Thanks for the suggestion.

### 2016-08-19 writeup/RochatEtal_ResponseToReviewers.tex

We thank reviewer \#1 for the speedy review. Below we address the reviewer’s
comments and suggestions.

\begin{enumerate}

\item {\it In the bottom paragraph on P.10, the authors emphasized that "At scales larger
than 20km, the April and October spectra based on hourly velocity snapshots are
nearly indistinguishable from each other". This emphasis appears to be contradictory
to the statements in other places of the manuscript that the seasonality of subinertial
submesoscale flows is strong (e.g., those relating to Figure 1b-c and Figure 2).}\\

    \bdp
        We agree that the emphasis on the similarity of hourly spectra can appear
        out of contradictory if placed out of context. Those spectra (now Figure 4c
        and supporting information) are similar because both submesoscale sub-inertial
        and super-inertial flows undergo a strong seasonal cycle that is out of
        phase --- there is a phase cancellation and the spectra based on hourly
        spectra is indistinguishable within errorbars. In any event, we removed
        that sentence and we believe the new Figure 4 will convey
        the phase cancellation idea more efficiently.
    \edp

@@

We thank reviewer \#2 for the speedy review. Below we address the reviewer’s
comments and suggestions.

\subsection{Major points}

\begin{enumerate}

  \item {\it First of all, generation mechanism of submesoscale inertial-gravity waves in summer
        and early fall is unclear to me. In the paper, authors attribute the mechanism to
        internal tides (X-12, section 6, last sentence) and emphasize using the tidal forcing in
        their model (X-5, section 2, 2nd paragraph). However, authors do not explain or show
        any interaction mechanism of internal tides (please at least add some reference). With
        the fixed frequencies of tidal constituents and existence of the critical latitude for the
        diurnal tides, I am not sure what mechanism can represent the surface near-inertial
        wave generation within the whole analyzed region by internal tides (with the
        horizontal scales of 10-100km). If authors compare Figures 2 and 4 with those
        estimated from the non-tidal forcing run in the Supplemental material, authors'
        conclusion could be strengthened.}\\

        \bdp
          The generation mechanism generation of internal
          waves are changes in the wind (NIWs) and barotropic tides sloshing over
          topographic features --- we are not claiming that there exists a different
          generation mechanism
          in summer compared to winter; neither are we claiming that surface NIWs
          are generated by internal tides. What is different is the near-surface
          expression of IGWs, which is very sensitive to the near-surface stratification
          (D'saro 1978). This is particular true for higher vertical-mode waves
          (supplemental material).\\

          To reference potential mechanisms of how energy can quickly be transferred to
          high wavenumbers by interaction with mesoscale-submesoscale turbulence and
           PSI,  we added new references to the discussion (Ponte \& Klein 2015;
          Mackinon \& Winters 2004; Alford et al 2016). \\

          The simulation not forced with tides (ECCO2 Adjoint) has much coarser
          horizontal resolution and barely resolves the submesoscales; its grid
          spacing is 18 km, while its effective resolution larger than 70 km.
      \edp

  \item {\it I also wonder why authors ignore wind-induced near-inertial waves (X-13, the first
        line). It is well studied that wind-induced near-inertial motion in mixed layer in back
        ground flow fields disperse quickly and propagate into the ocean interior (e.g., Young
        and Ben-Jelloul 1997, J. Mar. Res., 55, 735-66). The distribution of those waves
        depends on the distribution of vorticity (or Laplacian of vorticity) of background
        flows (Klein and Llewellyn Smith 2001, J. Mar. Res., 59, 697-723; Klein et al. 2004, Q.
        J. R. Meteorol. Soc., 130, 1153-1166). Therefore, those wind-induced near-inertial
        waves could have the horizontal scales of 10-100km. Note that those studies are based
        on geostrophically balanced or quasigeostrophic turbulent field. However, authors
        may need to study interactions between submeso dynamics regime (Rossby number
        and Richardson number are close to 1) and near-inertial waves in winter and spring
        cases where I guess that non-linear terms may play a role. I don't think that authors
        need to explain all internal wave dynamics in this GRL paper. I would rather suggest
        to remove all discussions related to submesoscale inertial-gravity waves and focus on
        describing differences between October and April.}\\

        \bdp
            We are now citing the recent review on NIWs by Alford et al. (2016) that contain
            a section dedicated to mesoscale-turbulence-NIW interactions, citing all the
            aforementioned papers. We believe the new discussion together with Figure 4
            and Figure S7 will help clear any ambiguity. Those figures unambiguously
            show that most of the IGWs with horizontal scales between 10-100 km
            are mostly super-inertial (not near-inertial); the refracted/dispersed  NIWs
            propagate into the interior and only weakly project onto SSH.
        \edp

  \item {\it Finally, I don't understand why authors define 10-100 km as a submesoscale and used
        this scale for the spatial filtering (X-7, section 3). Authors should discuss Figure 4 first
        and then give a reason why they define 10-100 km as a submesoscale and use this
        scale for the filtering. In X-11, section 5, author point out that a part of the dynamics
        around 250 km scale should be explained by near-inertial waves. Then, I wonder why
        authors decide to discuss internal waves in 10-100 km scale.}

        \bdp
            Following reviewer's \#3 suggestion, we have added paragraphs about SWOT
            to the introduction and conclusion. This adds context to the manuscript
            and help justify why we focus only on 10-100 km scales. Our definition
            of submesoscales have pros and cons (see reply to reviewer's 3
            minor comment \#1) but is consistent with the definition used by
            (at least part of) the SWOT community (Fu\& Ferrari 2008). Giving the
            definition, which we present in the first sentence of the introduction,
            then there is no ambiguity in using 100 km as a cut-off scale for the filter.
            We prefer to present bulk statistics first, and then discuss details of
            different scales --- this makes even more sense now with the new wavenumber-frequency
            analysis.
        \edp

@@

\end{enumerate}

\subsection{Minor points}

\begin{enumerate}

  \item {\it X-4, section 1, 2nd paragraph and X-13, section 6 last: Authors claimed that they
        analyze the Kuroshio Extension region. However, their study region does not
        include between $~$140E and $~$155E where the Kuroshio Extension is strongest in
        Figure 1(a). Please explain why you exclude this region in the paper. Since those
        longitudes includes the Izu-Ogasawara ridge (internal tide generation) and
        formation region of Subtropical mode water (strong mixed layer front), I would
        expect that including these latitudes would strengthen authors conclusions. I
        would also like to point out that positions of KEO and KESS moorings in the
        supplemental material are outside of the study region.}\\

        \bdp
             The upper bound on
             the size of the domain is chosen for lateral homogeneity purposes,
             necessary for the spectral analysis. We decided to focus in
             this subdomain to avoid the strong lateral inhomogeneities associated
             with the Kuroshio upstream and near its separation from Japan's coast.
             We have consider larger and smaller domains and found insiginificant
             quantitative differences.\\

             The supporting information explicitly states that both KEO and KESS
             mooring were outside our chosen domain. We analyzed data from those
             mooring simply to assess the skill of numerical model in representing
             high-frequency modes. As discussed in the supporting information, the
             comparison uses model output in the grid points closest
             to the mooring.
        \edp

  \item {\it X-4, section 1, 3rd paragraph: “submesoscale inertial-gravity waves”. I feel that this
        nomenclature is too sudden. It might be better to write “inertial-gravity waves
        with a horizontal scales of 10-100 km (hereafter submesoscale inertial-gravity
        waves)”?}\\

        \bdp
          Definition included --- thanks for pointing that out.
        \edp

@@

  \item {\it X-9, section 4, 2nd paragraph: I wonder what we could learn from the sentence “In
          other words, even in April, when submesoscale turbulence prevails, only the
          daily-averaged fields are largely in geostrophic balance”. This suggests that the
          balanced and upward cascade motion (originated from the submesoscale
          instability) scenario in Sasaki et al., (2014, nature comm. 5) could be wrong
          because they use the daily averaged data (X-11, section 5, last of 1st paragraph)?
          Or it just means that the balanced regime is actually the submeso-dynamics
          regime?}

          \bdp
              This simply means that there is significant ageostrophic components,
              even in April. Of course, daily-averaging the fields filters most of
              ageostrophic. This does not imply that Sasaki et al's mechanism is
              incorrect; it does suggests, however, a more complicated picture, with
              likely small energy leak towards small scales. In any event, to avoid ambiguity, we
              have rewritten that sentence and excluded repeated information.
          \edp

  \item {\it X-10, section 4, 2nd paragraph, the last sentence: Please add some reference.}
          The sentence has been rewritten.
\end{enumerate}

### 2016-08-19 writeup/rocha_etal.tex
Adds many responses to reviewer 2

Submesoscale IGWs, however, peak in late summer/early fall,
when the upper ocean is strongly stratified. Thus there exists a
strong seasonal modulation of upper-ocean submesoscale dynamics:
submesoscale turbulence dominates the upper-ocean dynamics in late winter/early fall,
whereas submesoscale IGWs prevail in late summer/early fall.
Because submesoscale turbulence is weakest in late summer/early fall,
the present results indicate that IGWs account for most of the
summertime submesoscale sea-surface height variability with implications for SWOT.

@@

The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
 (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 joint-PDF is similar for hourly and daily-averaged fields, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results are characteristic of winter, indicating that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence.
 The hourly velocity and sea-surface heigh fields have an important
ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
winter (Figure \ref{fig3}e). In other words, even in April, when submesoscale turbulence
prevails, only the daily-averaged fields
are largely
in  geostrophic balance (the joint-PDF of daily-averaged vorticity vs.\\ Laplacian of sea-surface height
is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; Figure \ref{fig3}f).

@@

%
%
%\begin{figure*}[ht]
%  \begin{center}
%    \includegraphics[width=.65\textwidth]{figs/fig5.pdf}
% \caption{\small Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
% in the 1/48$^\circ$ simulation. Solid lines
% are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
% fields.}
% \label{fig5}
% \end{center}
%\end{figure*}
%
%
%Figure \ref{fig5}a depicts the horizontal wavenumber spectra of surface KE in
%April and October. At scales larger than 20 km, the April and October spectra based on hourly
%velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
%Consistent with the results of
%\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
%there is significant high-frequency variabilty at submesoscales. Daily-averaging
%the velocity field suppresses spatial variability at scales smaller than about 250
%km, both in April and October (compare solid lines against dashed lines in Figure
%\ref{fig5}a). But this suppression is dramatic in October, when the IGWs peak. At scales
%smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
%flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
%is strong, consistent
%with the results of \citet{sasaki_etal2014}, which are
%based on daily-averaged velocity fields of a different model without tidal forcing
%(P. Klein, personal communication).
%
%The IGWs project on the sea-surface.
%There is a dramatic difference between the spectra
%based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig5}b): at scales smaller than 100 km, the spectra
%of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
%sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
%flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
%turbulence and near-surface submesoscale IGWs conspire to yield weak seasonality
%in the spectra of KE and sea-surface height variance based on hourly fields.
%
%
\section{Discussion and conclusion}
Our work adds to recent modeling \citep{sasaki_etal2014}
and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
near-surface submesoscale IGWs in the Kuroshio Extension undergo
a strong seasonal cycle that is out of phase with the seasonal cycle of
submesoscale turbulence. We conjecture that the summertime dominance of IGWs
 \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
enhancement of IGWs due to re-stratification of
the upper ocean.

>>>>

IGWs with horizontal scales between 10-100 km (hereafter submesoscale IGWs),
however, peak in late summer/early fall,
when the upper ocean is strongly stratified. Thus there exists a
strong seasonal modulation of upper-ocean submesoscale dynamics:
submesoscale turbulence dominates the upper-ocean dynamics in late winter/early fall,
whereas submesoscale IGWs prevail in late summer/early fall.
Because submesoscale turbulence is weakest in late summer/early fall,
the present results indicate that IGWs account for most of the
summertime submesoscale sea-surface height variability with implications for SWOT.

@@

The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
 (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
 with strong positive skewness are fingerprints of submesoscale fronts
 \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
 joint-PDF is similar for hourly and daily-averaged fields, although
 the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
 The April results are characteristic of winter, indicating that wintertime submesoscale
 surface velocity is strongly dominated by submesoscale turbulence.
 The hourly velocity and sea-surface heigh fields have an important
ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
winter (Figure \ref{fig3}e). The daily-averaged fields
are largely
in  geostrophic balance (Figure \ref{fig3}f).

@@

%
%
%\begin{figure*}[ht]
%  \begin{center}
%    \includegraphics[width=.65\textwidth]{figs/fig5.pdf}
% \caption{\small Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
% in the 1/48$^\circ$ simulation. Solid lines
% are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
% fields.}
% \label{fig5}
% \end{center}
%\end{figure*}
%
%
%Figure \ref{fig5}a depicts the horizontal wavenumber spectra of surface KE in
%April and October. At scales larger than 20 km, the April and October spectra based on hourly
%velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
%Consistent with the results of
%\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
%there is significant high-frequency variabilty at submesoscales. Daily-averaging
%the velocity field suppresses spatial variability at scales smaller than about 250
%km, both in April and October (compare solid lines against dashed lines in Figure
%\ref{fig5}a). But this suppression is dramatic in October, when the IGWs peak. At scales
%smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
%flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
%is strong, consistent
%with the results of \citet{sasaki_etal2014}, which are
%based on daily-averaged velocity fields of a different model without tidal forcing
%(P. Klein, personal communication).
%
%The IGWs project on the sea-surface.
%There is a dramatic difference between the spectra
%based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig5}b): at scales smaller than 100 km, the spectra
%of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
%sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
%flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
%turbulence and near-surface submesoscale IGWs conspire to yield weak seasonality
%in the spectra of KE and sea-surface height variance based on hourly fields.
%
%
\section{Discussion and conclusion}
Our work adds to recent modeling \citep{sasaki_etal2014}
and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
near-surface submesoscale IGWs in the Kuroshio Extension undergo
a strong seasonal cycle that is out of phase with the seasonal cycle of
submesoscale turbulence. IGW energy and SSH variance are significantly spread across
the submesoscales, likely through mesoscale refraction and dispersion \citep[e.g.,][and references there in]{ponte_klein2015,alford_etal2016}
 and parametric
subharmonic instability \citep[e.g., ][]{mackinnon_winters2005}. In the LLC simultions,
super-inertial IGWs account for most of the surface submesoscale KE and SSH variance;
near-inertial IGWs project onto larger scales.

 %We further conjecture that the summertime dominance of IGWs
 %\citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
%enhancement of IGWs due to re-stratification of
%the upper ocean.
